Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-10-30 main.tex
Combined the eigenvalue plots for both gains into one panel plot.

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use the gain factor \(\kappa=8\) (low) and \(\kappa=10\) (high) to tune the gain
magnitude during the experiments. Scaling the proportional feedback gain
linearly with respect to speed stabilizes the normally unstable weave mode of
the bicycle down to 2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-gain-8-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=8\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-gain-10-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=10\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

>>>>

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use the gain factor \(\kappa=8\) (low) and \(\kappa=10\) (high) to tune the gain
magnitude during the experiments. Scaling the proportional feedback gain
linearly with respect to speed stabilizes the normally unstable weave mode of
the bicycle down to 2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=8\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}
